---
title: ''
output: pdf_document
geometry: margin=.75 in
header-includes:
   - \usepackage{cancel}
   - \usepackage{multirow,setspace}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{mathtools}
   - \usepackage{amsfonts}
   - \usepackage{amssymb}
---



```{r}
rm(list = ls())
setwd("G:\\math\\640")
options(scipen=999)
source("multiplot.R")
require(VGAM, quietly=T); require(invgamma, quietly=T); require(mvtnorm, quietly=T);
require(ggplot2, quietly=T); require(RColorBrewer, quietly=T); require(zoo, quietly=T)
wind <- read.table("wind.txt",header = F)

```



\begin{align*}
f(w_i) &= \cfrac{w_i}{\theta^2} \ \exp\left(-\cfrac{w_i^2}{2\theta^2}\right) \\[1.5 em]
\mathcal{L}(W|\theta^2) & \propto \prod_{i=1}^n \cfrac{1}{\theta^2} \  \exp\left(-\cfrac{w_i^2}{2\theta^2}\right) \\[.25 em]
&= \left(\theta^2\right)^{-n} \ \exp\left(-\cfrac{1}{2\theta^2} \sum_{i=1}^n w_i^2\right) \\[.25 em]
\end{align*}

A conjugate prior for $\theta^2$ is the inverse gamma
\begin{align*}
    \pi\left(\theta^2\right) & \propto \left(\theta^2\right)^{-\left(\alpha_0+1\right)} \ \exp\left( - \cfrac{\beta_0}{\theta^2} \right).
\end{align*}

And if we let $\alpha_0 = \beta_0 = 0$, we get an improper, non-informative flat prior
\begin{align*}
    \pi\left(\theta^2\right) & \propto \left(\theta^2\right)^{-1}. 
\end{align*}

Which leads to the posterior
\begin{align*}
    P(\theta^2 | W ) &\propto \left( \theta^2 \right)^{n+1}  \ \exp\left(-\cfrac{1}{2\theta^2} \sum_{i=1}^n w_i^2\right).
\end{align*}

And  
\begin{align*}
    \theta^2|W &\sim IG\left( n, \ \cfrac{1}{2 } \sum_{i=1}^n w_i^2   \right).
\end{align*}


\ 

Or we could transform $\theta^2 = \mu^{-2}$

\begin{align*}
\mathcal{L}(W|\mu^{-2}) & \propto \prod_{i=1}^n  \mu^2 \  \exp\left(-\mu^2 \cfrac{w_i^2}{2 }\right) \\[.25 em]
&= \left(\mu^2\right)^{n} \ \exp\left(-\cfrac{\mu^2}{2} \sum_{i=1}^n w_i^2\right) \\[1.425 em]
\ell (W|\mu^{-2}) & \propto n \log\left(\mu^2\right) - \cfrac{\mu^2}{2} \sum_{i=1}^n w_i^2 \\[.5 em]
\cfrac{\partial \ell}{\partial \left(\mu^2\right)} &\propto \cfrac{n}{\mu^2} - \cfrac{1}{2} \sum_{i=1}^n w_i^2 \\[.5 em]
\cfrac{\partial^2 \ell}{\partial \left(\mu^2\right)^2} &\propto -\cfrac{n}{\left(\mu^2\right)^2} \\[0.75 em]
\end{align*}
\begin{align*}
-E\left[\cfrac{\partial^2 \ell}{\partial \left(\mu^2\right)^2}\right] &= \left(\mu^2\right)^{-2}\\[.25 em]
\left[J\left(\mu^2\right)\right]^{\frac{1}{2}} &= \left(\mu^2\right)^{-1}
\end{align*}

\begin{align*}
    P\left( \mu^2 | W \right) & \propto  \left(\mu^2\right)^{n-1} \ \exp\left(-\cfrac{\mu^2}{2} \sum_{i=1}^n w_i^2\right) \\[1 em]
    \mu^2 | W & \sim Gamma\left( n, \ \cfrac{1}{2 } \sum_{i=1}^n w_i^2   \right)
\end{align*}


\ 

```{r, cache=T}
n <- nrow(wind)
Nsim <- 2000
set.seed(17)
thetatheta <- rinvgamma(Nsim , n+1 , (1/2)* sum(wind[,1]^2)   ) 
quantile(thetatheta , probs = c(0.025, 0.5, 0.975))

#mode
theta <- sqrt( thetatheta  )
summary(theta)
quantile(theta, probs = c(0.025, 0.5, 0.975))

qu <- quantile(theta , probs = c(0.5, 0.025, 0.975))[2:3]
d <- data.frame( density(theta)[[1]], round(density(theta)[[2]],6) )
colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]

ggplot( data = d , aes(x=x , y=y)    ) + geom_line( col="red"  ) +
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ), fill="red", alpha = .15) +
	theme(legend.position="none", axis.title.x=element_blank(), axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ theme_minimal()
 
rmean <- theta * sqrt(  pi / 2 )

summary(rmean )
quantile(rmean, probs = c(0.025, 0.5, 0.975))

qu <- quantile(rmean, probs = c(0.5, 0.025, 0.975))[2:3]
d <- data.frame( density(rmean)[[1]], round(density(rmean)[[2]],6) )
colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]

ggplot( data = d , aes(x=x , y=y)    ) + geom_line( col="blue"  ) +
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ), fill="blue", alpha = .15) +
	theme(legend.position="none", axis.title.x=element_blank(), axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ theme_minimal()

rmedian <- theta * sqrt(  2 * log(2) )
summary(rmedian )

summary(rmedian )
quantile(rmedian, probs = c(0.025, 0.5, 0.975))

qu <- quantile(rmedian, probs = c(0.5, 0.025, 0.975))[2:3]
d <- data.frame( density(rmedian)[[1]], round(density(rmedian)[[2]],6) )
colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]

ggplot( data = d , aes(x=x , y=y)    ) + geom_line( col="green"  ) +
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ), fill="green", alpha = .15) +
	theme(legend.position="none", axis.title.x=element_blank(), axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ theme_minimal()

```

\ 

```{r, cache=T}
PPD <- thetatheta *0
set.seed(17)
for( i in 1:Nsim){  PPD[i] <- rrayleigh(1,  sqrt(thetatheta[i])) }

qu <- quantile(PPD, probs = c(0.5, 0.025, 0.975))[2:3]
d <- data.frame( density(PPD)[[1]], round(density(PPD)[[2]],6) )
colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]

ggplot( data = d , aes(x=x , y=y)    ) + geom_line( col="purple"  ) +
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ), fill="purple", alpha = .15) +
	theme(legend.position="none", axis.title.x=element_blank(), axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ theme_minimal()+
   ggtitle("Sampled Posterior Density")
 
 
sum(1*(PPD > 15))/length(PPD)
```


\newpage 


The first model will be the casual users $y_c$. We first check to see if we should transform the outcome variable.

\ 


```{r, fig.width=6.5, fig.height=9.5}
cb <- read.table("day.txt",header = T)

par(mfrow=c(4,2) ,  mar=c(2.1,4.1,2.1,2.1) )
ty <-  (  cb$casual ); plot(density(ty), main="non-transformed y"); qqnorm(ty); qqline(ty )
ty <- scale(cb$casual) ; plot(density(ty), main="[y-mean(y)]/sd(y)"); qqnorm(ty); qqline(ty )
ty <- sqrt(  cb$casual ); plot(density(ty), main="sqrt(y)"); qqnorm(ty); qqline(ty )
ty <- log(  cb$casual ); plot(density(ty), main="log(y)"); qqnorm(ty); qqline(ty )
```

\ 

The first row of plots shows that $y_c$ should probably be transformed to ensure normality. Of the three transformation choices presented here, it appears $\sqrt{y_c}$ provides a reasonable assumption of normality. Using example I.22 we find the following results.

\ 

Now assume the transformed variable $y'_c$ has the distribution 
 
 \ 
 
 \hfil \(\displaystyle y'_c  \sim N(X \bm{\beta}, \sigma^2 I_n) \)

\ 

The likelihood is then
\begin{align*}
\mathcal{L}\left( y | X, \pmb{\beta}, \sigma^2 \right) & \propto \left|\sigma^2 I_n\right|^{-\frac{1}{2}} \ \exp\left[ -\cfrac{1}{2}
  \left(Y-X\pmb{\beta} \right)^T \left(\sigma^2 I_n\right) \left(y'_c -X\pmb{\beta} \right) \right] \\[.5 em]
  & = \left( \sigma^2 \right)^{-\frac{n}{2}}  \ \exp\left[ -\cfrac{1}{2 \sigma^2 } \left(y'_c -X\pmb{\beta} \right)^T   \left(y'_c-X\bm{\beta} \right) \right] 
\end{align*}

With the joint prior

 \hfil $\pi\left( \pmb{\beta} ,\sigma^2 \right)  \propto N(X \pmb{\beta} , \sigma^2 I_n)$

 \ 

which gives a joint posterior of 

\ 

 \hfil $P\left( \pmb{\beta}, \sigma^2 |  y , X  \right) \propto \left( \sigma^2 \right)^{-\frac{n}{2}-1}  \ \exp\left[ -\cfrac{1}{2 \sigma^2 } \left(y'_c -X\bm{\beta} \right)^T   \left(y'_c -X \pmb{\beta} \right) \right]$


\ 


We find 

\hfil $\pmb{\beta} | \sigma^2 \sim N\left[\hat{\pmb{\beta}}, \sigma^2 \left[X^T X\right)^{-1} \right]$ 

 \ 
 
and

\ 

\hfil $\sigma^2 | X \sim IG\left[\cfrac{n-k}{2},\cfrac{1}{2} \left(y'_c  - X  \hat{\pmb{\beta}} \right)^T\left(y'_c- X  \hat{\pmb{\beta}} \right) \right]$.

\ 

Now we can perform the analysis.

\ 

```{r, cache = T}
n <- nrow(cb )
Nsim <- 20000
y1 <- sqrt( cb[,1] )	#casual
X <- as.matrix( cbind(1, cb[,-(1:2)] ) )
colnames(X)	<- c("(Intercept)",colnames(cb[,-(1:2)] ))
p <- ncol(X)

bhat <-  (solve(t(X)%*%X)%*%(t(X)%*%y1))
#fit <- ( lm( y1~X[,-1]  ));
SSY <- SSYc <- t(y1 - X%*%bhat)%*%(y1 - X%*%bhat); #anova(fit)
#vcov( fit )
rbetas <- matrix(0, nrow = Nsim, ncol = p)
XtXi	<- solve(t(X)%*%X)

set.seed(13)
rsig	<- rinvgamma(Nsim , (n-p)/2, (1/2)*SSY)
for(i in 1:Nsim){
	CovX	<- rsig[i]*XtXi
	rbetas[i,] <- c(rmvnorm(1, mean = bhat, sigma = CovX ) )
}

ycrbetas <- rbetas
rbMat	<- apply(rbetas, 2, quantile, probs = c(0.5, 0.025, 0.975))
colnames(rbMat)	<- colnames(X) #coef(fit)
rbMat <- as.data.frame( round(t(rbMat), 4) )
rbMat$zero <- rbMat[,2] < 0 & rbMat[,3] > 0
rbMat

rbsum <- ( apply( rbetas , 2 , summary) )
colnames(rbsum )	<- colnames(X)
 rbsum <- t( rbsum ); rbsum
```

```{r, cache = T, fig.width=6.5, fig.height=9.5}
colnames(rbetas)	<- colnames(X)
plotz <- list()
for( j in 1:ncol(X)) {
	Tdat <- rbetas[,j]
qu <- quantile(Tdat, probs = c(0.5, 0.025, 0.975))[2:3]
d <- data.frame( density(Tdat)[[1]], round(density(Tdat)[[2]],6) )
colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]
J <- ifelse( j == 2 , 9,j ); COLR <- brewer.pal(9,"Set3")[J]

 plotz[[j]] <- ggplot( data = d , aes(x=x , y=y)    ) + 
		geom_line( col=COLR  ) + theme_minimal()+
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ),
		 fill=COLR , alpha = .15) + ylab("") +
	theme(legend.position="none", axis.title.x=element_blank(), 
		axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ 
   	ggtitle( colnames(rbetas)[j] ) 
 }


multiplot(plotz[[1]], plotz[[2]], plotz[[3]], plotz[[4]],plotz[[5]], plotz[[6]], plotz[[7]], plotz[[8]], cols=2)
```

\ 
```{r, eval=F}
## Y 
for( i in 1:Nsim) {Yc[  i,] <- rmvnorm(1,  (X %*%rbetas[i,])^2 , sigma = rsig[1] * diag(730) )}
```

```{r, echo = F}
Yc <- apply(ycrbetas, 1 , function(B) (X%*%B)^2)
```


```{r, cache = T}
## Y 
ycplots <- list()
ycsums <- list()

DATES <-  seq( as.Date("2011-01-01"), as.Date("2012-12-31"), by="+1 day")
DATES <-  DATES[-which( DATES  == "2012-02-29") ]
rownames(Yc) <- gsub(" ","",as.character(as.yearqtr(DATES)))

for( i in 1:length( unique(rownames(Yc)) ) ){
 	
	yq <- unique(rownames(Yc))[i]
 	yq <-  Yc[ which( rownames(Yc) == yq ) , ] 
	yq <-  colSums(yq) / dim(yq)[1] 
	ycsums[[i]] <- summary(yq)
	
	qu <- quantile(yq, probs = c(0.5, 0.025, 0.975))[2:3]
	d <- data.frame( density(yq)[[1]], round(density(yq)[[2]],6) )
	colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]
	J <- ifelse( i == 6 , 9,i ); COLR <- brewer.pal(9,"Set3")[J]
	COLR <- brewer.pal(9,"Set1")[J]
 

 ycplots[[i]] <- ggplot( data = d , aes(x=x , y=y)    ) + 
		geom_line( col=COLR  ) + theme_minimal()+
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ),
		 fill=COLR , alpha = .15) + ylab("") +
	theme(legend.position="none", axis.title.x=element_blank(), 
		axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ 
   	ggtitle(  unique(rownames(Yc))[i] ) 
 }

ycsums <- as.matrix(do.call(rbind, ycsums))
ycsums <- as.data.frame(t(ycsums))
colnames(ycsums) <- paste0("Yc",unique(rownames(Yc)))
```


\ 


We will conduct a similar analysis on registered users $y_r$.



\ 

```{r, cache = T, fig.width=6.5, fig.height=3.5}
par(mfrow=c(1,2) ,  mar=c(2.1,4.1,2.1,2.1) )
ty <-  (  cb[,2] ); plot(density(ty), main="non-transformed y"); qqnorm(ty); qqline(ty )
```

```{r, cache = T}
n <- nrow(cb )
Nsim <- 20000
y1 <- ( cb[,2] )	
X <- as.matrix( cbind(1, cb[,-(1:2)] ) )
colnames(X)	<- c("(Intercept)",colnames(cb[,-(1:2)] ))
p <- ncol(X)

bhat <-  (solve(t(X)%*%X)%*%(t(X)%*%y1))
#fit <- ( lm( y1~X[,-1]  )); coef(fit)
SSY <- SSYr <- t(y1 - X%*%bhat)%*%(y1 - X%*%bhat); #anova(fit)
#vcov( fit )
rbetas <- matrix(0, nrow = Nsim, ncol = p)
XtXi	<- solve(t(X)%*%X)

set.seed(235)
rsig	<- rinvgamma(Nsim , (n-p)/2, (1/2)*SSY)
for(i in 1:Nsim){
	CovX	<- rsig[i]*XtXi
	rbetas[i,] <- c(rmvnorm(1, mean = bhat, sigma = CovX ) )
}

yrrbetas <- rbetas
rbMat	<- apply(rbetas, 2, quantile, probs = c(0.5, 0.025, 0.975))
colnames(rbMat)	<- colnames(X) #coef(fit)
rbMat <- as.data.frame( round(t(rbMat), 4) )
rbMat$zero <- rbMat[,2] < 0 & rbMat[,3] > 0
rbMat

rbsum <- ( apply( rbetas , 2 , summary) )
colnames(rbsum )	<- colnames(X)
 rbsum <- t( rbsum ); rbsum
```

```{r, cache = T, fig.width=6.5, fig.height=9.5}
colnames(rbetas)	<- colnames(X)
plotzz <- list()
for( j in 1:ncol(X)) {
	Tdat <- rbetas[,j]
qu <- quantile(Tdat, probs = c(0.5, 0.025, 0.975))[2:3]
d <- data.frame( density(Tdat)[[1]], round(density(Tdat)[[2]],6) )
colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]
J <- ifelse( j == 2 , 9,j ); COLR <- brewer.pal(9,"Set3")[J]

 plotzz[[j]] <- ggplot( data = d , aes(x=x , y=y)    ) + 
		geom_line( col=COLR  ) + theme_minimal()+
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ),
		 fill=COLR , alpha = .15) + ylab("") +
	theme(legend.position="none", axis.title.x=element_blank(), 
		axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ 
   	ggtitle( colnames(rbetas)[j] ) 
 }


multiplot(plotzz[[1]], plotzz[[2]], plotzz[[3]], plotzz[[4]],plotzz[[5]], plotzz[[6]], plotzz[[7]], plotzz[[8]], cols=2)
```

\ 

```{r, eval=F}
## Y 
for( i in 1:Nsim) {Yr[  i,] <- rmvnorm(1,  (X %*%rbetas[i,])^2 , sigma = rsig[1] * diag(730) )}
```

```{r, echo = F}
Yr <- apply(yrrbetas, 1 , function(B) (X%*%B))
```

```{r, cache = T}
yrplots <- list()
yrsums <- list()

DATES <-  seq( as.Date("2011-01-01"), as.Date("2012-12-31"), by="+1 day")
DATES <-  DATES[-which( DATES  == "2012-02-29") ]
rownames(Yr) <- gsub(" ","",as.character(as.yearqtr(DATES)))

for( i in 1:length( unique(rownames(Yr)) ) ){
 	
	yq <- unique(rownames(Yr))[i]
 	yq <-  Yr[ which( rownames(Yr) == yq ) , ] 
	yq <-  colSums(yq) / dim(yq)[1] 
	yrsums[[i]] <- summary(yq)
	 
	qu <- quantile(yq, probs = c(0.5, 0.025, 0.975))[2:3]
	d <- data.frame( density(yq)[[1]], round(density(yq)[[2]],6) )
	colnames(d) = c("x" , "y" ); d$area <- d[,1] > qu[1] & d[,1] < qu[2]
	J <- ifelse( i == 6 , 9,i ); COLR <- brewer.pal(9,"Set3")[J]
	COLR <- brewer.pal(9,"Set1")[J]
 

 yrplots[[i]] <- ggplot( data = d , aes(x=x , y=y)    ) + 
		geom_line( col=COLR  ) + theme_minimal()+
	geom_ribbon(data = d[which(d$area == T),], aes(x, ymin=0 , ymax=y ),
		 fill=COLR , alpha = .15) + ylab("") +
	theme(legend.position="none", axis.title.x=element_blank(), 
		axis.title.y=element_blank(),
	     plot.title = element_text(hjust = 0.5)   )+ 
   	ggtitle(  unique(rownames(Yc))[i] ) 
 }


yrsums <- as.matrix(do.call(rbind, yrsums))
yrsums <- as.data.frame(t(yrsums))
colnames(yrsums) <- paste0("Yc",unique(rownames(Yr)))
```

\ 

We can now compare the two outcome variables for the model. For simplicity, I have aggregated the in-model predictions. We know have their means by quarter. In other words, $X\hat{\pmb{\beta}}$ is an $730 \times 20000$ matrix. I have condensed the days into quarters by taking the mean count of users for each quarter. So we have the distribution of the mean (20000 observations) for the eight quarters in 2011-2012. It should be noted that the user count for casual riders is $\left(X\hat{\pmb{\beta}}\right)^2$ because of the square root transformation on $y_c$.

The casual user's predictions will always come first. Here are the summaries:

\

```{r}
ycsums
yrsums
```

\ 

There are many more registered riders than there are casual riders. They have similar jumps and falls in ridership as it gets warmer/colder. It is interesting to note that the jump for casual riders in Q1 to Q3 is about triple or quadruple. The jump for registered users in the same time period is less than double. 

\ 

Also it is clear that Capital Bikeshare increased in popularity from 2011 to 2012. It appears ridership almost doubles.

\ 


```{r, cache = T, fig.width=6.5, fig.height=4.25}
multiplot(ycplots[[1]], ycplots[[2]], ycplots[[3]], ycplots[[4]],ycplots[[5]], ycplots[[6]], ycplots[[7]], ycplots[[8]], cols=2)
```
```{r, cache = T, fig.width=6.5, fig.height=4.25}
multiplot(yrplots[[1]], yrplots[[2]], yrplots[[3]], yrplots[[4]],yrplots[[5]], yrplots[[6]], yrplots[[7]], yrplots[[8]], cols=2)
```



\ 

Prediction (from Gelman pg. 357):

```{r, cache=T}
pred <- c(1,1,0,1,.344348,0.34847,0.804783,0.179117)

ycp <- ycpreds <- apply( ycrbetas , 1 , function(B) (pred %*% B)^2 )
ycpreds <- ycpreds*0

set.seed(2011)
rsig	<- rinvgamma(Nsim , (n-p)/2, (1/2)*SSYc)
for(i in 1:Nsim){
	ycpreds[i] <- (rnorm(1, mean = ycp[i], sqrt(rsig[i])  ) )
}

summary(ycpreds);quantile(ycpreds, probs = c(0.5, 0.025, 0.975))

yrp <-  yrpreds <- apply( yrrbetas , 1 , function(B) pred %*% B )
yrpreds <- yrpreds*0

set.seed(2012)
rsig	<- rinvgamma(Nsim , (n-p)/2, (1/2)*SSYr)
for(i in 1:Nsim){
	yrpreds[i] <- c(rnorm(1, mean = yrp[i], sqrt(rsig[i])  ) )
}

summary(yrpreds); quantile(yrpreds, probs = c(0.5, 0.025, 0.975))



```

